Evaluating infectious disease forecasts with allocation scoring rules

### 2023-11-06 alloscore_manuscript/alloscore-application.Rnw

Infectious disease forecasting models have emerged as important tools in public health. Predictions of disease dynamics have been used to inform decision making about a wide variety of measures to reduce disease spread and/or mitigate the severity of disease outcomes. For example, estimates of expected onset of flu season have been used to aid national vaccination strategies \citep{igboh2023timing}, and forecasts of Ebola dynamics have been used to allocate surveillance resources \citep{meltzer2014estimating, rainisch2015regional}. \cite{bertsimas2021predictionsCOVID} developed tools to inform decision making from infectious disease forecasts, which have been used to inform allocation of limited medical supplies such as ventilators, ICU capacity planning, and vaccine distribution strategy. Models developed by \cite{fox_real-time_2022} have been used to inform resource and care site planning, as well as community guidelines for masking, traveling, dining and shopping (University of Texas, 2022)\nocite{utnews2022}. In April of 2022, the Centers for Disease Control and Prevention (CDC) announced the launch of the Center for Forecasting and Outbreak Analytics (CFA) to translate disease forecasts into decision making (CDC, 2022)\nocite{cdc2022cfa}, indicating that this has been identified as an important direction at the highest levels of government public health response. The value of infectious disease forecasts has typically been measured by how closely they predict disease outcomes such as cases, hospitalizations or deaths using, for example, root mean square error (RMSE) \citep{papastefanopoulos2020covid} or weighted interval score (WIS) \citep{bracher2021evaluating}, however, recently authors have been calling for evaluating forecasts through their impact on policy \citep{marshall2023predictions}.

@@

However, we are aware of only a limited body of work that explicitly attempts to measure the value of infectious disease forecasts through their impact on policy, and much of this discussion has proceeded informally. For example, \cite{ioannidis2022forecastingCOVIDfailed} discuss the possible negative consequences of inaccurate forecasts of infectious disease, but do not attempt to quantify the utility or loss incurred as a result of those forecasts. Separately, there is a thread of literature that does quantify the link between infectious disease modeling and policy making, but this work has been done outside of a forecasting context. As an example, \cite{Probert2016decisionMakingFootMouth} develop measures of the cost of actions designed to control a hypothetical outbreak of foot-and-mouth disease and use this framework to explore policy recommendations from a variety of simulation-based projection models.

@@

We see this work as an initial overture for what we hope will grow to be a large, collaborative body of work in more closely coupling applied epidemiological forecasting with public health decision-making.
We note that others have explored similar linkages.
For example, \cite{bilinski_adaptive_2023} explore ways in which policymaker preferences could inform risk thresholds for predictive models.
\ngr{Would be nice to have a few other examples here.}
In some situations, individual or ensemble models could be developed to minimize the loss for a particular setting, although this is also an area largely unexplored in the realm of public health to date, with some initial methodological development in econometrics.\citep{loaiza-maya_focused_2021}

>>>>

Infectious disease forecasting models have emerged as important tools in public health. Predictions of disease dynamics have been used to inform decision making about a wide variety of measures to reduce disease spread and/or mitigate the severity of disease outcomes. For example, estimates of expected onset of flu season have been used to aid national vaccination strategies \citep{igboh2023timing}, and forecasts of Ebola dynamics have been used to allocate surveillance resources \citep{meltzer2014estimating, rainisch2015regional}. \cite{bertsimas2021predictionsCOVID} developed tools to inform decision making from infectious disease forecasts, which have been used to inform allocation of limited medical supplies such as ventilators, ICU capacity planning, and vaccine distribution strategy. Models developed by \cite{fox_real-time_2022} have been used to inform resource and care site planning, as well as community guidelines for masking, traveling, dining and shopping (University of Texas, 2022)\nocite{utnews2022}. In April of 2022, the Centers for Disease Control and Prevention (CDC) announced the launch of the Center for Forecasting and Outbreak Analytics (CFA) to translate disease forecasts into decision making (CDC, 2022)\nocite{cdc2022cfa}, indicating that this has been identified as an important direction at the highest levels of government public health response. The value of infectious disease forecasts has typically been measured by how closely they predict disease outcomes such as cases, hospitalizations or deaths using, for example, root mean square error (RMSE) \citep{papastefanopoulos2020covid} or weighted interval score (WIS) \citep{bracher2021evaluating}.  However, recently authors have been calling for evaluating forecasts through their impact on policy \citep{marshall2023predictions, bilinski_adaptive_2023}.

@@

However, we are aware of only a limited body of work that explicitly attempts to measure the value of infectious disease forecasts through their impact on policy, and much of this discussion has proceeded informally. For example, \cite{ioannidis2022forecastingCOVIDfailed} discuss the possible negative consequences of inaccurate forecasts of infectious disease, but do not attempt to quantify the utility or loss incurred as a result of those forecasts. \cite{bilinski_adaptive_2023} explore ways in which policymaker preferences could inform risk thresholds for predictive models using a framework for measuring the costs and losses associated with taking an action that is similar to methods that have been used in meteorology and elsewhere. \cite{marshall2023predictions} develop a forecast scoring rule that is informally motivated by utility considerations, but the score is not derived from a decision-theoretic set up. Separately, there is a thread of literature that quantifies the link between infectious disease modeling and policy making outside of a forecasting context. As an example, \cite{Probert2016decisionMakingFootMouth} develop measures of the cost of actions designed to control a hypothetical outbreak of foot-and-mouth disease and use this framework to explore policy recommendations from a variety of simulation-based projection models.

@@

We see this work as an initial overture for what we hope will grow to be a large, collaborative body of work in more closely coupling applied epidemiological forecasting with public health decision-making.
We note that a few papers have begun exploring similar linkages as described in the literature review \textendash but we see much room for additional work in this area.
In some situations, individual or ensemble models could be developed to optimize scores that are attunend to a particular decision making setting.
This area is also largely unexplored in the realm of public health to date, with some initial methodological development in econometrics \citep{loaiza-maya_focused_2021}.
